---
title: "Getting Started with R"
author: "David Bell"
output: 
  html_document:
    keep_md: TRUE
    code_folding: show
    theme: readable
---

### R Script vs. R Markdown

Data scientists traditionally use two files in RStudio: R scripts (.R) and R Markdown documents (.rmd). When we're exploring, cleaning, and visualizing data, we'll use an R script where we don't have to worry about making our work neat or presentable. But when it comes time to present our findings, R Markdown becomes a great tool to write to and make an HTML document where we can create a more cohesive analysis. You can create either of these by selecing `File`, `New File`, then select whichever format you need (notice the numerous coding language options available to you in RStudio).

### Loading Libraries

First things first, we must load our required libraries. Click `Insert`, `R` to add an R chunk. A good standard is always loading the `tidyverse` package as it contains many of the tools and functions we'll need to wrangle and visualize data. We do this by typing `library(tidyverse)`.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

If a library you're wishing to use hasn't yet been installed, you will need to run `install.packages()` with whichever library you're loading in quotes. The function I usually use is `pacman::p_load()` which loads the library, and if it hasn't yet been installed, attempts to install it. It's a bit of a combination of `install.packages()` and `library()`.

### Exploring Data

To view a data frame in another tab, run `View(data)`. Let's take a look at the `mtcars` dataset:

```{r}
# View(mtcars)
```

This is a good way to sort explore the data, where you can sort the data by individual columns. Another good command is `summary(data)`:

```{r}
summary(mtcars)
```
If your data is quantitative, this outputs a six-number summary for each variable. If you want a brief glance of your data, try `head(data, 10)` which displays the first ten rows of your data frame:

```{r}
head(mtcars, 10)
```
`str(data)` will list the structure of your data frame:

```{r}
str(mtcars)
```

It's our job as data scientists to be literate about our/our client's data and this means exploring and getting familiar with the data. Checking for missing values is part of that. We can check by running `is.na(data)` and `sum(is.na(data))`:

```{r}
is.na(mtcars)
sum(is.na(mtcars))
```

`is.na(data)` isn't always the best option if you're working with large data, but the `sum(is.na(data))` can give you an idea of how much you're missing. 

### Using the Pipe Function (%>%)

Since R is a statistical/data driven language, most functions require the data set on which you're performing the function. An example is `select(mtcars, mpg)` which selects the `mpg` column from `mtcars`.

```{r}
select(mtcars, mpg)
```
Alternatively, we could use the pipe function. The pipe function is a way to perform a sequence of operations. How would we use this example? Well, instead of feeding the `select()` function `mtcars`, we could pipe `mtcars` *into* the `select` function.

```{r}
mtcars %>% 
  select(mpg)
```
What if we wanted to do multiple operations on `mtcars`? We could use another pipe.

```{r}
mtcars %>% 
  select(mpg) %>% 
  summary()
```
The pipe command becomes a great way to link together operations that would have otherwise been performed separately. The shortcut is `Cmd Shift M` on Mac, and `Ctrl Shift M` on PC. We'll be using the pipe statement consistently throughout bootcamp.




