---
title: "Reading in Data and Using the Assignment Operator"
author: "David Bell"
output: 
  html_document:
    keep_md: TRUE
    code_folding: show
    theme: readable
---

# Reading in Data

Before we start working with a dataset, we need to import it into our R Studio environment. The format your data is in will determine the manner in which you read it in. Today we'll be working with the `Iris` data set. Before we do anything, we need to read in our required libraries.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
```

Here's a link to the Iris .csv dataset: `https://gist.githubusercontent.com/curran/a08a1080b88344b0c8a7/raw/0e7a9b0a5d22642a06d3d5b9bcbad9890c8ee534/iris.csv` Let's go ahead and import it using the `read_csv()` function in `readr`. We do this by typing `read_csv()` and then the file path to the dataset - in this case, the url above. Don't forget to put the file path in quotes!

```{r echo=TRUE, message=FALSE, warning=FALSE}
read_csv("https://gist.githubusercontent.com/curran/a08a1080b88344b0c8a7/raw/0e7a9b0a5d22642a06d3d5b9bcbad9890c8ee534/iris.csv")
```

We see that `read_csv()` did the trick in reading in our data, but we aren't yet able to manipulate it. First we need save it as tibble (an R data frame). We do this using the assignment operator. 

## Using the Assignment Operator

In other programming languages such as Java and Python, if we want to assign something to an object name, we use the `=` assignment operator. But in R, the assignment operator is `<-`. The shortcut is `Option + -` on Mac and `Alt + -` on PC. Let's call our dataset `dat_iris` using the assignment operator.

```{r echo=TRUE, message=FALSE, warning=FALSE}
dat_iris <- read_csv("https://gist.githubusercontent.com/curran/a08a1080b88344b0c8a7/raw/0e7a9b0a5d22642a06d3d5b9bcbad9890c8ee534/iris.csv")
```

Now our data is in an R tibble that we can manipulate as we please.

## Exploring the Data

Remember what we learned last time. As soon as we start working with a dataset, we should explore and get familiar with the data. Go ahead and run some exploratory commands (hint: `head()`, `str()`, `summary()`).

```{r echo=TRUE, message=FALSE, warning=FALSE}
head(dat_iris)
summary(dat_iris)
str(dat_iris)
```


## Next Time

Next time we'll work on creating some visualizations using tools in the `GGPlot2` library. 

 








