---
title: "Data Manipulation Using `dplyr`"
author: "David Bell"
output: 
  html_document:
    keep_md: TRUE
    code_folding: show
    theme: readable
---

## What is `dplyr`?

`dplyr` is  package within the `tidyverse` which contains some handy data manipulation functions. We'll spend our time this week learning about and using some of these functions. Here's some `dplyr` [documentation](https://dplyr.tidyverse.org/) if you'd like to do some more reading. 

## `select()`

The first function we'll look at is `select()`. This function allows us to grab a particular column (or columns) a data frame. Here's a few ways we could use `select()`.

Say we want to get a five-number summary from only the `mpg` and `disp` columns in the `mtcars` dataset. 

```{r echo = TRUE, message = FALSE, warning = FALSE}
library(tidyverse)
```

```{r echo=TRUE, warning=FALSE}
mtcars %>%
  select(c(mpg, disp)) %>% 
  summary()
```

`select()` allowed me to grab only the columns I wanted to run the `summary()` function on. Also notice that since I'm selecting multiple columns, I need to use the `c()` command which indicates I'm passing a list into the function.

Another instance where we often use `select()` is when creating a new data frame that's a subset of our data. Say I want a new data frame that's only the `mpg`, `disp`, and `cyl` columns.

```{r}
mtcars_subset <- mtcars %>% 
  select(c(mpg, disp, cyl))
```

I now have a new data frame that's only those three columns. I can now do other operations and commands on that data frame the same way I could with `mtcars`.

## `filter()`

The `filter()` function selects instances within a data frame based on certain criteria. This criteria is usually given using logical expressions. Let's look at some examples. 

Say I wanted to get a five-number summary for the `disp` column, but only for cars that have six cylinders. 

```{r}
mtcars %>% 
  filter(cyl == 6) %>% 
  select(disp) %>% 
  summary()
```

The conditional statment `==` is one of equality. Say I wanted to get the same summary, but for all cars *except* those with six cylinders.

```{r}
mtcars %>% 
  filter(cyl != 6) %>% 
  select(disp) %>% 
  summary()
```

`!=` is saying select every row, but leave out the ones where `cyl = 6`. `filter()` also accepts greater than (`>`), less than (`<`), greater than or equal to (`>=`), less than or equal to (`<=`), within a given range (`between()`), or near a certain value (`near()`).

## `mutate()`

If I wanted to create a new column in my data frame, I would use `mutate()`. This allows new columns to be created based on existing columns. Say I wanted a new column which is the ratio of `mpg` and `hp`.

```{r}
mtcars <- mtcars %>% 
  mutate(mpg_hp_ratio = mpg / hp)

colnames(mtcars)
```

I now have a new column called `mpg_hp_ratio` which is just `mpg` divided by `hp`. I can do all sorts of operations in `mutate()`.

```{r}
mtcars <- mtcars %>% 
  mutate(average_disp = mean(disp))

colnames(mtcars)
summary(mtcars$average_disp)
```

The sky's the limit with `mutate()`.

## `arrange()`

If you're ever looking to sort or order your data by a certain column, use `arrange()`.

```{r}
mtcars %>% 
  arrange(disp) %>% 
  head()
```

This returns the first six rows of `mtcars`, in ascending order of the `disp` column (least to greatest). If you wanted the data displayed in *descending* order (greatest to least), you can use the `desc()` function.

```{r}
mtcars %>% 
  arrange(desc(disp)) %>% 
  head()
```





